---
title: "Assortative Mating Model"
author: "Dan Conroy-Beam"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview
 
### Purpose

Determine whether assortative mating can emerge from individual mating motivations. Agents will select one another as mates based on a physical attractiveness variable.

Each agent will pursue mates probabilistically and in proportion to mates' physical attractiveness. The model will then calculate the extent to which agents mate assortatively for attractiveness.

For comparison, the model will also assess assortative mating for randomly formed pairs.

This model is based loosely on Kalick and Hamilton (1986).
 
### Agents

Each agent will have two randomly assigned features: sex and physical attractiveness. Half of all agents will be male and half will be female.

Agent physical attractiveness values will be drawn from random uniform distributions constrained between 0 and 10.

Agents will also possess a maximum number of dates they will tolerate going on before settling for any mate.

Agent selectivity will gradually decline as they approach their maximum number of dates.

### Life Cycle

1. Date
2. Offer
3. Pair
 
#### Date

At the start of the life cycle, a random agent will be paired with a random opposite-sex partner. Dating agents' number of dates will be incremented by one.
 
#### Offer

Next, each dating agent will decide whether to make a long-term commitment offer to their dating partner. The probability of making an offer will be proportional to their date's attractiveness, adjusted by the number of dates the agent has gone on.

If an agent has reached their maximum number of dates, they will make an offer to their next date regardless of attractiveness.

#### Pair

Finally, dating agents will pair if both agents mutually make long-term commitment offers to one another.
 
The life cycle will repeat until all agents are paired
 
### Analysis

The model will compute the correlation between male and female attractiveness within couples for each model run. Correlations across model runs will be plotted with a histogram.

The model will also randomly scramble agent couples and compute the attractiveness correlation for random pairs within model run. 

These random choice correlations will be plotted alongside the attractiveness-driven results for comparison.



## Model



### Packages
```{r packages}

library(ggplot2)

```



### Parameters

```{r parameters}

#Population Size#
#Total number of agents to be generated
popSize<-100



#Maximum Number of Dates#
#Maximum number of dates agents will go on before giving up
maxDates<-50



#Model Loops#
#Number of times to repeat the model
modelLoops<-500

```



### Functions

```{r functions}

#Agent Generation#
#A function to generate agents
agentGenerate<-function(n,sex){
  
  #n = number of agents to generate
  #sex = sex of desired agents
  
  #Assign agents a random physical attractiveness
  physAtt<-runif(n,0,10)
  
  #Create a vector to store the number of dates agents have been on
  numDates<-0
  
  #Put the agents together
  agents<-data.frame(sex,physAtt,numDates)
  
  #Output the agents
  return(agents)
}

```



### Life Cycle

```{r lifeCycle}

#Generate a blank dataframe to store model results
results<-data.frame("loop"=1:modelLoops,
                    "type"=rep(c("random","nonRandom"),each=modelLoops),
                    "r"=NA)

#Repeat the model modelLoops times
for (m in 1:modelLoops) {
  
  #Generate the agents
  females <- agentGenerate(popSize / 2, "female")
  males <- agentGenerate(popSize / 2, "male")
  
  #Create empty vectors to store paired agents
  pairedFemales <- c()
  pairedMales <- c()
  
  
  while (length(pairedMales) < nrow(males)) {
    #Date#
    
    #Create a vector of all agents
    mDate <- 1:nrow(males)
    fDate <- 1:nrow(females)
    
    #Limit this vector to single agents
    mDate <- mDate[!(mDate %in% pairedMales)]
    fDate <- fDate[!(fDate %in% pairedFemales)]
    
    
    #If more than one male remains...
    if (length(mDate) > 1) {
      #Pick one at random to go on a date
      mDate <- sample(mDate, 1)
    }
    
    #Same thing for females
    if (length(fDate) > 1) {
      fDate <- sample(fDate, 1)
    }
    
    #For the agents going on a date,
    #increment their number of dates by 1
    females$numDates[fDate] <- females$numDates[fDate] + 1
    males$numDates[mDate] <- males$numDates[mDate] + 1
    
    
    
    #Offer#
    
    #Calculate how selective each agent should be based on number of dates completed so far
    fChoosiness <- (maxDates - females$numDates[fDate]) / maxDates
    mChoosiness <- (maxDates - males$numDates[mDate]) / maxDates
    
    #Prevent choosiness values less than 0
    fChoosiness <- ifelse(fChoosiness < 0, 0, fChoosiness)
    mChoosiness <- ifelse(mChoosiness < 0, 0, mChoosiness)
    
    #Set the probability of making an offer proportional to (1) date attractiveness and (2) number of dates so far
    fProb <- ((males$physAtt[mDate] ^ 3) / 1000) ^ fChoosiness
    mProb <- ((females$physAtt[fDate] ^ 3) / 1000) ^ mChoosiness
    
    #Have each agent probabilistically make an offer of commitment
    fOffer <- rbinom(1, 1, fProb)
    mOffer <- rbinom(1, 1, mProb)
    
    
    
    #Pair#
    
    #If both agents made mutual offers...
    if (fOffer & mOffer) {
      #Add their index to the paired agent vectors
      pairedFemales <- c(pairedFemales, fDate)
      pairedMales <- c(pairedMales, mDate)
    }
    
  }
  
  #Store the attractiveness correlations in the results dataframe for both non-random and random partner choice
  results$r[results$loop==m & 
              results$type=="nonRandom"]<-cor(females$physAtt[pairedFemales], 
                                              males$physAtt[pairedMales])
  
    results$r[results$loop==m & 
                results$type=="random"]<-cor(sample(females$physAtt),
                                             sample(males$physAtt))
    
}


```



### Analysis

```{r analysis}

p1<-ggplot(data=results,
           aes(x=r,
               fill=type))+
  labs(x="Correlation",y="Density",fill="Choice Type")+
  geom_density(alpha=.5)+
  scale_fill_manual(labels=c("Non-Random","Random"),
                    values=c("dodgerblue3","forestgreen"))+
  coord_cartesian(xlim=c(-1,1))

p1

```

