---
title: "Infectious Disease"
author: "Dan Conroy-Beam"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview
 
### Purpose

A simple model of infectious disease transmission. Agents will move around a 2-dimensional surface and potentially spread an infectious disease to those nearby.

The model will manipulate the number of agents who move vs. stay in place (i.e. social distance) and observe the effect of social distancing on disease spread.
 
### Agents

A population of 200 agents will be generated. Agents will have a random position on a 2D surface.

Each agent will have an infection status (either susceptible, infected, or recovered); all but one agent will start the model as susceptible. One agent will start as infected.

All agents will also have a variable that tracks the number of days they have been infected; this will start at 0 for uninfected agents and 1 for infected agents.

Finally, all agents will possess a variable that reflects whether they are social distancing or not.

The proportion of agents that are social distancing will be set by a social distancing parameter.
 
### Life Cycle
 
1. Move
2. Infect
3. Recover

#### Move

Agents will move in a random direction. The distance they move on each day of the life cycle will be set by a movement speed parameter.

Agents that are social distancing will not move
 
#### Infect

Infected agents will potentially transmit the disease to susceptible agents nearby them. 

Only agents within a certain radius of an infected agent are at risk of transmission; 

This radius will be set as a parameter.

The probability of infecting each agent within this radius will also be set by a transmission probability parameter.

Only susceptible agents can be infected; infected and recovered agents cannot be re-infected
 
#### Recovery

The infection duration will increment by 1 each day an agent is infected. 

When the total duration exceeds an infection duration parameter, the agent will change status to recovered
 
### Analysis

The model will save the proportion of the population that is susceptible, infected, and recovered each day

These will be plotted as a function of time to observe both the rate of disease spread and the peak infected proportion as a function of the social distancing parameter.


## Model

### Packages

```{r packages}

library(ggplot2)

```



### Parameters

```{r parameters}

#Population Size#
#Number of agents to generate
popSize<-200



#Social Distancing#
#Proportion of agents who are social distancing
socialDist<-.9



#Movement Speed#
#How far agents move in a given time step
moveSpeed<-12



#Infection Radius#
#How close two agents must be to transmit disease
radius<-6


#Transmission Probability#
#The probability an infected agent will pass the disease to a susceptible agent
contagion<-.5



#Recovery Time#
#How long agents take to recover from infection
recoveryTime<-14



#Number of Days#
#Number of "days" to repeat the life cycle
numDays<-100

```



### Functions

```{r functions}

#Agent Generation#
#A function to generate agents
agentGenerate<-function(n,s){
  
  #n = population size
  #s = socialDist; whether agents are social distancing
  
  #Assign each agent a random position on a 2-D surface
  xPos<-runif(n,0,100)
  yPos<-runif(n,0,100)
  
  #Assign each agent a social distancing status
  distancing<-sample(c(rep("moving",n-floor((n*s))),
    rep("standing",floor(n*s))))
  
  #Assign each agent a disease status
  status<-sample(c(rep("susceptible",n-1),"infected"))
  
  #Create a variable to track duration of infection
  daysInfected<-ifelse(status=="infected",1,0)
  
  #Put the agents together
  agents<-data.frame(xPos,yPos,distancing,status,daysInfected)
  
  #OUtput the agents
  return(agents)
  
}



#Move#
#A function to make agents move
move<-function(startPos,speed){
  
  #Generate a random degree of squared movement on the x-dimension
  xDiff<-runif(1,0,speed^2)
  
  #Allocate any remaining movement speed to the y-dimension
  yDiff<-speed^2-xDiff
  
  #Square root to convert into coordinate dimensions
  xDiff<-sqrt(xDiff)
  yDiff<-sqrt(yDiff)
  
  #Randomly make each movement direction positive or negative
  moveDiff<-sample(c(-1,1),2,replace=T)*c(xDiff,yDiff)
  
  #Generate new coordinates
  newPos<-startPos+moveDiff
  
  #If necessary, wrap agent position back around to a valid position in the 2-D surface
  newPos<-newPos %% 100
  
  #Output the new coordinate position
  return(newPos)
  
}



#Infect#
#Spread the infection between agents
infect<-function(agents,radius,contagion){
  
  #Agents = all agents
  #Radius = exposure radius
  #Contagion = infection probability per exposure
  
  #Generate a distance matrix
  distances<-as.matrix(dist(agents[,1:2]))
  
  #Determine which agents are currently infected
  infected<-which(agents$status=="infected")
  
  #Loop through infected agents...
  for(sickPerson in infected){
    
    #Determines which agents are within the infection radius of the sick person
    transmit<-distances[sickPerson,]<=radius
    
    #Determine which exposed agents are actually susceptible
    transmit<-transmit & agents$status=="susceptible"
    
    #Determines whether a transmission event will actually happen
    transmit<-transmit & rbinom(nrow(agents),1,contagion)
    
    #Change the infection status of people who experienced a transmission event
    agents$status[transmit]<-"infected"
    
  }
  
  return(agents)
  
}

```



### Life Cycle

```{r lifeCycle}

#Generate the agents
agents<-agentGenerate(popSize,socialDist)



###Life Cycle###

#Generate a blank vector for storing results
results<-data.frame("day"=rep(1:numDays,each=3),
                    "status"=c("susceptible","infected","recovered"),
                    "p"=NA
                    )

#Loop through days...
for (day in 1:numDays) {
  
  #Move#
  
  #Loop agent-by-agent...
  for (a in 1:nrow(agents)) {
    #If the agent is not social distancing...
    if (agents$distancing[a] == "moving") {
      #Have them move to a new position
      agents[a, 1:2] <- move(agents[a, 1:2], moveSpeed)
      
    }
    
  }
  
  
  
  #Infect#
  
  #Determine which agents get infected today
  agents <- infect(agents, radius, contagion)
  
  
  
  #Recovery#
  
  #Increments days infected for infected agents
  agents$daysInfected[agents$status == "infected"] <-
    agents$daysInfected[agents$status == "infected"] + 1
  
  #If agents have been infected for at least recoveryTime days, change their status to recovered
  agents$status <- ifelse(agents$daysInfected >= recoveryTime,
                          "recovered",
                          agents$status)
  
  #Record proportion of agents who have each infection status as of this day
  results$p[results$day==day & 
              results$status=="susceptible"] <- mean(agents$status=="susceptible")
  
    results$p[results$day==day & 
              results$status=="infected"] <- mean(agents$status=="infected")
    
      results$p[results$day==day & 
              results$status=="recovered"] <- mean(agents$status=="recovered")
  
}


```



### Analysis

```{r analysis}

sirPlot<-ggplot(data=results,
                aes(x=day,
                    y=p,
                    color=status)
                )+
  labs(x="Time (Days)",
       y="Proportion of Population",
       color="Infection Status")+
  scale_color_manual(values=c("red",
                              "forestgreen",
                              "goldenrod"),
                     labels=c("Infected",
                              "Recovered",
                              "Susceptible"))+
  geom_line(lwd=2)

sirPlot

```

