---
title: "Life-Dinner Principle Model"
author: "Dan Conroy-Beam"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview
 
### Purpose

Observe the "life-dinner principle": in coevolutionary arms races, prey animals have the advantage in that they are fighting for their lives whereas predators are just fighting for dinner.

Agents representing rabbits and foxes will have random encounters with one another. Whether the rabbit escapes--and whether the fox eats--will depend on their relative speed.

Rabbits will reproduce only if they survive; foxes will reproduce only if they get enough to eat.
 
### Agents

Agents will represent either foxes or rabbits. Each agent will also have a speed variable randomly assigned to them from a normal distribution. Rabbit agents will possess a variable that determines whether they have been eaten; foxes possess a variable that tracks the number of rabbits they have eaten.
 
### Life Cycle
 
1. Encounter
2. Eat/escape
3. Reproduce/Die
 
#### Encounter

During this life stage, each fox will encounter one rabbit at random. 
 
#### Eat/Escape

The model will next determine whether the fox eats its encountered rabbit or whether the rabbit escapes. The model will roll dice for each agent with the maximum value equal to the agent's speed value.
 
If the rabbit's dice roll is higher, the rabbit escapes; otherwise the fox eats the rabbit. 

The total number of encounters a fox experiences will be set by a parameter.
 
#### Reproduce/Die

After encounters are resolved, the agents will reproduce. Rabbits will reproduce only if they survived. Foxes will reproduce in proportion to the number of rabbits they ate.

Offspring will inherit their speed variable from their parents.

### Analysis

The model will compute the average speed of foxes and rabbits within each model run.



## Model

### Packages
```{r packages}

library(ggplot2)

```



### Parameters
```{r parameters}

#Model Loops#
#Number of times to repeat the full model
modelLoops<-10



#Fox Population Size#
#Number of fox agents to generate
foxPop<-100



#Number of Fox-Rabbit Encounters#
#Number of rabbits each fox will encounter
nEncounters<-3



#Rabbit Population Size#
#Number of rabbit agents to generate
rabbitPop<-foxPop*nEncounters



#Selection Strength#
#Ratio of reproductive success between most and lease successful fox
selStrength<-.15



#Generations#
#Number of generations to simulation evolution
generations<-100

```



### Functions
```{r functions}

agentGenerate<-function(n,animal){
  
  #Assign each agent an ID number
  ID<-1:n
  
  #Assign a random speed to each agent
  speed<-rnorm(n,25,5)
  
  #If the agent is a rabbit...
  if(animal=="rabbit"){
    
    #Create a variable to track whether it has been eaten
    dead<-0
    
    #Put the rabbits together
    agents<-data.frame(ID,animal,speed,dead)
    
    #If the agent is a fox...
  } else {
    
    #Create a variable to track the number of rabbits each fox has eaten
    meals<-0
    
    #Put the foxes together
    agents<-data.frame(ID,animal,speed,meals)
    
  }
  
  #Output the agents
  return(agents)
  
}



#Chase#
#A function to simulate the chase between foxes and encountered rabbits
chase<-function(foxes,eRabbits){
  
  #Foxes = all foxes
  #eRabbits = encountered rabbits
  
  #Roll dice for rabbits and foxes based on their top speeds
  foxRolls<-runif(nrow(foxes),0,foxes$speed)
  rabbitRolls<-runif(nrow(eRabbits),0,eRabbits$speed)
  
  #Determine which foxes catch their rabbits
  catch<-foxRolls>rabbitRolls
  
  #Output the result of each encounter
  return(catch)
}



#Reproduce#
#A function to reproduce agents
reproduce<-function(agents,selStrength,n){
  
  #If the agent is a rabbit...
  if(unique(agents$animal)=="rabbit"){
    
    #Remove the agents that were eaten
    agents<-agents[agents$dead==0,]
    
    #Generate random offspring from the surviving parent rabbits
    offspring<-agents[sample(1:nrow(agents),n,replace=T),]
    
    #Reset offspring IDs
    offspring$ID<-1:n
    
  } else {
    
    #Remove foxes that did not catch any rabbits
    agents<-agents[agents$meals>0,]
    
    #Rescale meals to 0-to-1
    p<-agents$meals-min(agents$meals)
    p<-p/max(p)
    
    #Add a constant to probability weights to fix strength of selection
    p<-p+(1/selStrength)
    
    #Generate random fox offspring from the surviving parents
    offspring<-agents[sample(1:nrow(agents),
                             n,
                             replace=T,
                             prob=p),]
    
    #Reset offspring IDs
    offspring$ID<-1:n
    
    #Reset offspring meals
    offspring$meals<-0
    
  }
  
  #Output the offspring
  return(offspring)
  
}

```



### Life Cycle
```{r lifeCycle}

#Generate a blank dataframe to store model
results<-data.frame(
  "loop"=rep(1:modelLoops,each=generations*2),
  "generation"=1:generations,
  "animal"=rep(c("fox","rabbit"),each=generations),
  "speed"=NA
)


for (m in 1:modelLoops) {
  
  #Generate the agents
  foxes <- agentGenerate(foxPop, "fox")
  rabbits <- agentGenerate(rabbitPop, "rabbit")
  
  #Loop over generations of simulated evolution...
  for (g in 1:generations) {
    
    #Loop over encounters betwen foxes and rabbits...
    for (e in 1:nEncounters) {
      #Encounter#
      
      #Determine which rabbits will have an encounter with a fox this day
      encountered <- sample(rabbits$ID[rabbits$dead == 0], foxPop)
      
      #Eat/Escape#
      
      #Simulate encounteres between foxes and rabbits
      caught <- chase(foxes, rabbits[encountered, ])
      
      #Update whether each rabbit was eaten
      rabbits$dead[encountered] <- caught
      
      #Update the number of rabbits eatten by each fox
      foxes$meals <- foxes$meals + caught
      
    }
    
    #Reproduction#
    
    #Overwrite parent agents with offspring agents
    foxes <- reproduce(foxes, selStrength, foxPop)
    rabbits <- reproduce(rabbits, selStrength, rabbitPop)
    
    #Store the average speed of offspring agents
    results$speed[results$animal=="fox" &
                    results$loop==m & 
                    results$generation==g]<-mean(foxes$speed)
    
    results$speed[results$animal=="rabbit" &
                    results$loop==m &
                    results$generation==g]<-mean(rabbits$speed)
    
  }
  
}



```



### Analysis
```{r analysis}

p1<-ggplot(data=results,
           aes(x=generation,
               y=speed,
               color=animal))+
  labs(x="Generation",y="Animal Speed (Mean)",
       color="species")+
  stat_summary(fun=mean,geom="line")+
  scale_color_manual(labels=c("Fox","Rabbit"),
                     values=c("orangered","gray60"))

p1

```

